---
title: "Data exploration"
output:
  html_document:
    toc: true
    toc_float: true
  word_document: default
  pdf_document:
    toc: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r load packages, include=FALSE}
library(tidyverse)
library(sf)
library(readr)
library(knitr)
library(pROC)
library(gridExtra)
library(scales)
library(corrgram)
library(corrplot)
```

```{r setup, include=FALSE}
my_wd_all_dat <- "G:/ai_daten/P1070_USBW/MbE_MUC"

knitr::opts_knit$set(root.dir = normalizePath(my_wd_all_dat))
```

```{r read data, include=FALSE}
# dens_grid <- st_read("C:/Users/Vera/Documents/SUBDENSE/Projects/Liverpool_Dembski/R Output/grid_full.gpkg") %>% st_drop_geometry() #Pfad Vera
data_raw <- st_read("G:/ai_daten/P1047_SUBDENSE/liverpool_paper/01_data_input/in_vera/grid_full.gpkg")  #Pfad Denise


# par_with_dep_raw <- st_read("G:/ai_daten/P1070_USBW/MbE_MUC/workfiles_03_git/data_prep_output/parcels_10_final_data.gpkg") 
# var_desc <- read.csv("./workfiles_03_git/data_models/240925_MbE_MUC_data - Erklärende Variablen.csv")
```

# Data preparation
```{r}
#reduce to grid cells in built-up area 2011
data <- data_raw %>% filter(builtup2011 == 1) %>% 
  mutate(oac_challenged = as.numeric(ifelse(GRP %in% c("7a", "7b", "7c", "7d", "8a", "8b", "8c", "8d"),1,0)),     
         oac_students = as.numeric(ifelse(GRP %in% c("2a", "2b"),1,0)),     
         oac_success = as.numeric(ifelse(GRP %in% c("2c", "2d", "3d", "5a"),1,0))) %>% 
  st_drop_geometry()
```


# Explore explanatory variables

```{r overview explanatory var, include=FALSE}
# summary(data)
sum(is.na(data))
colSums(is.na(data))
names(data)
```

```{r functions, include=FALSE}

# functions numeric variables
print_description_variables <- function(var_desc, name_of_variable) {

kable(var_desc %>% 
        select(-var_name_german, -input_data, -reporting_date, -literature) %>% 
        filter(var_name == name_of_variable) %>% 
        pivot_longer(cols = category:rationale, names_to = "name", values_to = "value"), 
      col.names = c("", ""), 
      caption = "Basic information",
      format.args = list(decimal.mark = '.', big.mark = ","))
  
}

print_plots_numeric_var <- function(var_dat, name_of_variable) {

p1 <- var_dat %>%
  ggplot(aes(x = .data[[name_of_variable]]))+
  geom_histogram()+
  theme_light()

p2 <- var_dat %>%
  ggplot(aes(x = .data[[name_of_variable]]))+
  geom_boxplot()+
  theme_light()

grid.arrange(p1, p2, ncol = 2, heights = 1)

}

# functions categorical/logic variables
print_table_number_cat <- function(var_dat, name_of_variable) {
  
kable(var_dat %>% 
        group_by(.data[[name_of_variable]]) %>% 
        summarise(n = n()) %>% 
        arrange(desc(n)), 
      format.args = list(decimal.mark = '.', big.mark = ",")
      )
}

print_barplot_cat_logi <- function(var_dat, name_of_variable) {

var_dat %>% 
  ggplot(aes(y = .data[[name_of_variable]]))+
  geom_bar()+
    theme_light()
}
```

## Numeric variables

### amenity_count

```{r amenity_count, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
# print_description_variables(var_desc, "amenity_count")
summary(data$amenity_count)
print_plots_numeric_var(data, "amenity_count")
```

### m_to_park

```{r m_to_park, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
# print_description_variables(var_desc, "m_to_park")
summary(data$m_to_park)
print_plots_numeric_var(data, "m_to_park")
```

  
### m_to_train

```{r m_to_train, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
# print_description_variables(var_desc, "m_to_train")
summary(data$m_to_train)
print_plots_numeric_var(data, "m_to_train")
```

### min_to_livmain

```{r min_to_livmain, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "min_to_livmain")
summary(data$min_to_livmain)
print_plots_numeric_var(data, "min_to_livmain")
```


### nb11_LDcount

```{r nb11_LDcount, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "nb11_LDcount")
summary(data$nb11_LDcount)
print_plots_numeric_var(data, "nb11_LDcount")
```

### nb11_MDcount

```{r nb11_MDcount, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "nb11_MDcount")
summary(data$nb11_MDcount)
print_plots_numeric_var(data, "nb11_MDcount")
```


### nb11_HDcount

```{r nb11_HDcount, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "nb11_HDcount")
summary(data$nb11_HDcount)
print_plots_numeric_var(data, "nb11_HDcount")
```
         

### nb11_NBcount

```{r nb11_NBcount, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "nb11_NBcount")
summary(data$nb11_NBcount)
print_plots_numeric_var(data, "nb11_NBcount")
```

       
### sfh_share

```{r sfh_share, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "sfh_share")
summary(data$sfh_share)
print_plots_numeric_var(data, "sfh_share")
```

### Rank

```{r Rank, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "Rank")
summary(data$Rank)
print_plots_numeric_var(data, "Rank")
```
                   
### deprivation

```{r deprivation, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "deprivation")
summary(data$deprivation)
print_plots_numeric_var(data, "deprivation")
```

### income_rank

```{r income_rank, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "income_rank")
summary(data$income_rank)
print_plots_numeric_var(data, "income_rank")
```


### addresses_2013

```{r addresses_2013, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "addresses_2013")
summary(data$addresses_2013)
print_plots_numeric_var(data, "addresses_2013")
```

## Categorical variables

### dens_group11 

```{r dens_group11, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "dens_group11")
print_table_number_cat(data, "dens_group11")
print_barplot_cat_logi(data, "dens_group11")
```

                    


```{r LSOA11CD, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
### LSOA11CD
#print_description_variables(var_desc, "LSOA11CD")
# print_table_number_cat(data, "LSOA11CD")
# print_barplot_cat_logi(data, "LSOA11CD")
```

### dominant_year

```{r dominant_year, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
#print_description_variables(var_desc, "dominant_year")
print_table_number_cat(data, "dominant_year")
print_barplot_cat_logi(data, "dominant_year")
```



## Logical variables

### agriculture

nature           agriculture      industry         water            sports          
parks            port             airport          dump             rail

```{r agriculture, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
#print_description_variables(var_desc, "agriculture")
summary(data$agriculture)
print_barplot_cat_logi(data, "agriculture")
```

### oac_challenged

```{r oac_challenged, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
#print_description_variables(var_desc, "oac_challenged")
summary(data$oac_challenged)
print_barplot_cat_logi(data, "oac_challenged")
```

### oac_students

```{r oac_students, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
#print_description_variables(var_desc, "oac_students")
summary(data$oac_students)
print_barplot_cat_logi(data, "oac_students")
```

### oac_success

```{r oac_success, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
#print_description_variables(var_desc, "oac_success")
summary(data$oac_success)
print_barplot_cat_logi(data, "oac_success")
```


## Correlations between explanatory variables

```{r correlations all,  echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=6}
cor_var <- data %>%
  dplyr::select(amenity_count:oac_success) %>%
  dplyr::select(-builtup2011) %>% 
  dplyr::select(where(is.numeric)) %>%
  cor(use = "pairwise.complete.obs", method = "spearman")

corrplot(cor_var, type="lower",
         tl.cex = 0.6, cl.cex = 0.6,
         tl.col = "black")
corrplot(cor_var,
         type="lower",
         method = "number",
         tl.cex = 0.6, cl.cex = 0.6, number.cex = 0.5,
         tl.col = "black")
```

Correlations > 80

Correlations > 70


## Write data

```{r, include=FALSE}
#st_write(dat_bu_ldh_fin, "./workfiles_03_git/data_models/dat_all_bu.gpkg", delete_layer = TRUE)
```

